\documentclass{article}
\usepackage{ times}

\begin{document}

\begin{center}
Eric Gebhart
\linebreak
6908 Hillside Court
\linebreak
Raleigh, NC, 27607
\linebreak
Work: 531-5817
\linebreak
Home: 851-0176
\linebreak
\end{center}

\paragraph{Overview}
Over 21 years of experience as a systems developer.  Thirteen of those years
as a consultant.  The last eight years as a system developer at SAS Institute.       

\paragraph{10/95-Present: SAS Institute}
As a systems developer for SAS Institute designed and developed the ODS HTML,
and Markup destinations as well as the infrastructure used by those destinations
and others.  Created the first working version of the ODS RTF destination.  Created
all the original ODS styles.  

Wrote one paper on ODS Markup and have given several presentations on ODS Markup and ODS
in general.  As well as developing and teaching several classes on ODS Markup and tagsets.
Answer many questions about tagsets, as well as creating them to solve problems for
customers and SAS consultants.

Wrote a CVS wrapper in python that makes CVS more powerful and more aware of the SAS
programming environment.  Given presentations and written articles on using CVS
and the development environment at SAS.
Also contributed extensively towards making FreeBSD a viable
development platform at SAS.  Written several enhancements to the GNU Emacs editor to allow
more seamless integration with the SAS development environment.  Including an editing mode
for tagsets.

\paragraph{3/94-10/95: Consultant for Northern Telecom Inc.}          
Helped design and implement a trouble ticket system using JAM and JAM/PI.
Designed, developed and implemented an object oriented database manager to encapsulate a 
database using SQL with dynamic low-level binding.  Designed, developed and implemented an
object oriented, state driven screen manager, which controlled all screen functionality and interfaced automatically into the database manager.  All SQL statements were automatically generated by the interaction of these two tools.  Created code templates for the different types of screens which wolud exist in the application.  Taught a team of devolepers in the use of the templates and tools available.  These templates and tools allowed a realitively small team to develop a complex GUI in aa very short period of time.  Worked with two other projects which wished to interface their application with ours, in real time.  This was accomplished using sockets.  Designed and implemented a development environment, using RCS and GNU make.  This environment enabled simultaneous ports to different machines using one set of source code and makefiles.  This environment also allowed for
support of previous releases as well as multiple development branches.  Wrote tools to make the programming environment easier to use with ksh, awk, and perl.

\paragraph{11/92-3/94: Consultant for ABB.}
Designed an object oriented function library to facilitate the creation and maintenance of motif data entry screens.  Designed, developed and implemented an object oriented database manager to encapsulate a database using SQL with low-level dynamic binding.  Completely redesigned the underlying database shared by three different systems.  Ported the database manager to many different platforms and taught various development teams in it's use.  Designed and implemented a development environment using GNU Make.  This environment allowed simultaneous ports to various platforms using one set of source code and makefiles.  Helped design and implement an editor using JAM/PI which interfaced to an existing MOTIF system.  Wrote several UNIX Shell/AWK code generation programs.  Served as a consultant to other development teams for database design and software development methodologies and use of the database manager.  Also administered the Solaris Operating System and an Oracle 7 database on a Sun Workstation.  Served as a consultant for database design and software development methodologies for many software development teams at ABB.

\paragraph{7/92-10/92: Consultant for Carolina Mutual Wholesale Drug.}
Designed and implemented an object oriented library to handle all database and curses calls.
This library enabled quick and reliable development of a windowed interface to an informix
database.


\paragraph{1989-7/92: Consultant for Northern Telecom Inc.}
Helped design and create a database of over sixty tables.  Ported code
from HP, to DEC, Apollo, and Sun Workstations for bench marking purposes.
Installed and administered a Oracle database.  Designed and implemented functions which kept the
look and feel of the system consistent.  Encapsulated code which accessed JAM screens.  Designed
and created program templates and functions to expedite code
generation for a graphical user interface.  Developed a library of functions which encapsulated TBL and Nroff to simplify the creation of complex reports.  Developed many tools using shell scripts to make the programming environment easier to use.  Integrated many tools into the X Windows/motif environment.  Setup and maintained the boot files for all Xterminals in the system.

\paragraph{1986-1989: Inpedendent Intructor for various customers}
Taught Oracle database administration, SQL, UNIX system administration, Troff, C, Awk and shell
programming.  All on various platforms including; Sperry 5000/80, 5000/60, 5000/40, NCR, and HP-9000.

\paragraph{1986-1987: Instructor for Boeing Military Aircraft Co.}
Taught object orinted programming methods using C, Pascal and Modula2 on Apollo WorkStations.

\paragraph{1984-1989: Consultant for Independent Oil and Gas.}
Designed and implemented an Oracle Database and publishing system
to track and report on oil and gas wells.  Various reports were
generated daily using troff to render them for the printing press.  Subscribers
to the information service could also recieve index cards covering
specific areas.  These cards were also generated using Troff. Data
could also be bought covering specific township/ranges.  These data
files were automatically generated in the format required for
uploading into the customer's geographic applications.

\end{document}
